### 2021-06-24 thesis/tex/sterileTheory.tex
update

%\input{commands}
%\begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}.  
Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly, but saw the excess in The
$\ne$ events too. Together with the so-called reactor and gallium anomalies, these reports suggested 
that the three massive neutrino framework could be amended.
By introducing a fourth heavy mass state $\nu_4$, both appearance and disappeareance anomalies could be minimally accomodated.

However, we know from the decay width of the $Z$ boson that it only can interact with three flavor species\footnote{
    Direct measurements of the invisible $Z$ decay yields $N=2.92 \pm 0.05$~\cite{pdg}.
}, 
so this fourth mass state cannot be interacting weakly. In other words, it needs to transform as a
singlet under the broken gauge group $SU(2)_L \times U(1)_{EM}$.
We now distinguish between the three original neutrino flavors ($e$, $\mu$, and $\tau$) and the new fourth 
flavor by calling the former \emph{active} neutrinos
and the latter \emph{sterile} due to its non-interacting behavior.

@@

\subsection{The Hamiltonian}
The inclusion of the sterile state in the Hamiltonian is straightforward. We extend the PMNS matrix to incorporate the new flavor and mass eigenstates: 
\begin{align}
    U_{4gen} =
    \begin{pmatrix}
    U_{e1} & U_{e2} & U_{e3} & U_{e4} \\
    U_{\mu1} & U_{\mu2} & U_{\mu3} & U_{\mu4} \\
    U_{\tau1} & U_{\tau2} & U_{\tau3} & U_{\tau4} \\
    U_{s1} & U_{s2} & U_{s3} & U_{s4}
    \end{pmatrix}\,,
\end{align}
where a common parametrization of this new mixing matrix is 
\begin{align}\label{eq:U4gen_param}
    U_{4gen} = R_{34}R_{24}R_{14}U_{3gen}\,.
\end{align}
The mass matrix extends analogously:
\begin{align}\label{eq:dm4gen}
    M^2_{4gen} =
    \begin{pmatrix}
        0 & 0 & 0 & 0\\
        0 & \dm_{21} & 0  & 0 \\
        0 & 0 & \dm_{31} & 0 \\
        0 & 0 & 0 & \dm_{41} \\
    \end{pmatrix}\,.
\end{align}

Now, the interaction with matter requires a careful reconsideration of the matter potential. We start of with the unaltered potential matrix. 
Just as with the PMNS matrix, we extend this to $4\times4$:
\begin{align}\label{eq:Vsterile1}
    \begin{pmatrix}
        V_{CC} + V_{NC} & 0 & 0 & 0 \\
        0 &V_{NC} & 0 & 0 \\
        0 & 0 & V_{NC} & 0 \\
        0 & 0 & 0 & 0 
    \end{pmatrix}\,.
\end{align}
Now we need to include the terms that describes the matter potential felt by the sterile flavor state. Recalling our discussion above, we remind ourselves that the sterile neutrino by definition
does not participate in any interaction\footnote{The exception to this is of course gravity. The sterile neutrino is not massless.}. Thus, all potential terms involving the sterile state are zero. 
In other words, the potential matrix in Eq.~\ref{eq:Vsterile1} is complete, save for the usual subtraction by a constant identity matrix:

>>>>

%\input{commands}
%\begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}.  
Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly, but observed an excess in the
$\ne$ events too. Together with the so-called reactor and gallium anomalies, these reports suggested 
that the three massive neutrino framework could be amended.
By introducing a fourth heavy mass state $\nu_4$, both appearance and disappeareance anomalies could be minimally accomodated.
However, we know from the decay width of the $Z$ boson that it only can interact with three flavor species\footnote{
    Direct measurements of the invisible $Z$ decay yields $N=2.92 \pm 0.05$~\cite{pdg}.
}, 
so this fourth mass state cannot be interacting weakly. In other words, it needs to transform as a
singlet under the broken gauge group $SU(2)_L \times U(1)_{EM}$.
We now distinguish between the three original neutrino flavors ($e$, $\mu$, and $\tau$) and the new fourth 
flavor by calling the former \emph{active} neutrinos
and the latter \emph{sterile} due to its non-interacting behavior.

@@

\subsection{The Hamiltonian}
The inclusion of the sterile state in the Hamiltonian is straightforward. We extend the PMNS matrix
with terms for the mixing between the new flavor $(U_{si})$ and mass $(U_{\a 4})$ eigenstates:
\begin{align}
    U_{4gen} =
    \begin{pmatrix}
    U_{e1} & U_{e2} & U_{e3} & U_{e4} \\
    U_{\mu1} & U_{\mu2} & U_{\mu3} & U_{\mu4} \\
    U_{\tau1} & U_{\tau2} & U_{\tau3} & U_{\tau4} \\
    U_{s1} & U_{s2} & U_{s3} & U_{s4}
    \end{pmatrix}\,,
\end{align}
where a common parametrization of this new mixing matrix is 
\begin{align}\label{eq:U4gen_param}
    U_{4gen} = R_{34}R_{24}R_{14}U_{3gen}\,.
\end{align}
This parametrization does not affect the physics, but determines the structure of the mixing angles.

The mass matrix extends analogously:
\begin{align}\label{eq:dm4gen}
    M^2_{4gen} =
    \begin{pmatrix}
        0 & 0 & 0 & 0\\
        0 & \dm_{21} & 0  & 0 \\
        0 & 0 & \dm_{31} & 0 \\
        0 & 0 & 0 & \dm_{41} \\
    \end{pmatrix}\,.
\end{align}

Now, the interaction with matter requires a careful reconsideration of the matter potential. We start off with the unaltered potential matrix from Eq.~\ref{eq:V_matrix}. 
Just as with the PMNS matrix, we extend this to $4\times4$:
\begin{align}\label{eq:Vsterile1}
    \begin{pmatrix}
        V_{CC} + V_{NC} & 0 & 0 & 0 \\
        0 &V_{NC} & 0 & 0 \\
        0 & 0 & V_{NC} & 0 \\
        0 & 0 & 0 & 0 
    \end{pmatrix}\,.
\end{align}
Now we need to include the terms that describes the matter potential felt by the sterile flavor state. Recalling our discussion above, we remind ourselves that the sterile neutrino by definition
does not participate in any interaction\footnote{The exception to this is of course gravity. The sterile neutrino is not massless.}. Thus, all potential terms involving the sterile state are zero. 
In other words, the potential matrix in Eq.~\ref{eq:Vsterile1} is complete, save for the usual subtraction by a constant times the identity matrix:
